---
title: "Regression Models Course Project"
author: "Todd Kemmerling"
date: "August 28, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
data("mtcars")
```

## Questions To Be Answered

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”

"Quantify the MPG difference between automatic and manual transmissions"

## Grading Criteria

1. Did the student interpret the coefficients correctly?
2. Did the student do some exploratory data analyses?
3. Did the student fit multiple models and detail their strategy for model selection?
4. Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
5. Did the student do a residual plot and some diagnostics?
6. Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?
7. Was the report brief (about 2 pages long) for the main body of the report and no longer than 5 with supporting appendix of figures?
8. Did the report include an executive summary?
9. Was the report done in Rmd (knitr)?

Your report must be:

Written as a PDF printout of a compiled (using knitr) R markdown document.
Brief. 

Roughly the equivalent of 2 pages or less for the main text. 

Supporting figures in an appendix can be included up to 5 total pages including the 2 for the main report. The appendix can only include figures.

Include a first paragraph executive summary.

```{r mtcars}
summary(mtcars)
```

## Including Plots

You can also embed plots, for example:

```{r eda, echo=FALSE}
plot(mtcars$am, mtcars$mpg)
abline(lm(mtcars$mpg ~ mtcars$am))
```
```{r eda1, echo=FALSE}
mtcars$transtype <- 'Automatic'
mtcars[mtcars$am == 1, ]$transtype <- 'Manual'
mtcars$transtype <- as.factor(mtcars$transtype)
plot(mtcars$transtype, mtcars$mpg)
abline(lm(mtcars$mpg ~ mtcars$transtype))
```

```{r eda2, echo=FALSE}
plot(mtcars$am, mtcars$mpg)
abline(lm(mtcars$mpg ~ mtcars$am))
```

```{r eda3, echo=FALSE}
plot(mtcars$wt, mtcars$mpg)
abline(lm(mtcars$mpg ~ mtcars$wt))
```

```{r eda4, echo=FALSE}
plot(mtcars$disp, mtcars$mpg)
abline(lm(mtcars$mpg ~ mtcars$disp))
```

```{r eda5, echo=FALSE}
plot(mtcars$drat, mtcars$mpg)
abline(lm(mtcars$mpg ~ mtcars$drat))
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
