---
title: "Regression Models Course Project"
author: "Todd Kemmerling"
date: "August 31, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
data("mtcars")
mtcars$transtype <- 'Automatic'
mtcars[mtcars$am == 1, ]$transtype <- 'Manual'
mtcars$transtype <- as.factor(mtcars$transtype)
```

## Summary

This analysis is done from a sample of 32 cars of various makes and models. 19 of the cars have automatic transmissions and 13 have manual transmissions. This is a case where a simple look at the data can be a somewhat misleading. Constructing a simple linear model from the data appears to show a clear MPG advantage with a manual transmission. If we dig a bit deeper the weight of the vehicle also plays a significat part in MPG and at some vehicle weights the type of transmission makes no difference. It is not possible to conclude from this data that the transmissioin type makes any difference in the MPG of the vehicle.

## Details of the Analysis

```{r smodel}
fit <- lm(formula = mtcars$mpg ~ mtcars$transtype)
lift <- fit$coefficients[2]
```

The plot and model below clearly shows a benefit (or lift) of `r round(lift,2)` MPG by using a manual transmission.


```{r summaryplot, echo=FALSE}
plot(mtcars$transtype, mtcars$mpg , ylab = 'MPG', main='Transmission Type versus MPG')
abline(lm(mtcars$mpg ~ mtcars$transtype), col='blue')
```

The next plot shows two data sets, one for each transmission type. It shows the influence of weight on MPG for each transmission type. The following R code is used to determine the impact per 1000 pounds of increased weight.

```{r avmlifts}
liftAuto <- lm ( mtcars[mtcars$am == 0, ]$mpg ~ mtcars[mtcars$am == 0, ]$wt )$coefficients[2]
liftManual <- lm ( mtcars[mtcars$am == 1, ]$mpg ~ mtcars[mtcars$am == 1, ]$wt )$coefficients[2]
```

This shows for a manual transmission there is decrease of `r -round(liftManual,2)` MPG per 1000 pound increase weight. The corresponding decrease for an automatic transmission is `r -round(liftAuto,2)` MPG per 1000 pound increase weight.

```{r plot2, echo=FALSE}
plot ( mtcars[mtcars$am == 0, ]$wt, mtcars[mtcars$am == 0, ]$mpg , pch=0, col='blue', xlab = 'Weight x 1000 lbs', ylab = 'MPG', main = 'Weight vs Transmission vs MPG' )
abline(lm ( mtcars[mtcars$am == 0, ]$mpg ~ mtcars[mtcars$am == 0, ]$wt ), col='blue')
points ( mtcars[mtcars$am == 1, ]$wt, mtcars[mtcars$am == 1, ]$mpg , pch=1, col='red')
abline(lm ( mtcars[mtcars$am == 1, ]$mpg ~ mtcars[mtcars$am == 1, ]$wt ), col='red')
legend(x='topright', legend = c('Automatic','Manual'), col = c('blue','red'), pch= c(0,1) )
```

```{r finalmodel}
fit <- lm(mtcars$mpg ~ mtcars$transtype + mtcars$wt)
res <- residuals(fit)
fits <- fitted(fit)
plot(res ~ fits, main = 'Residuals at Fitted Points For Transmission + Weight', ylab = 'Residuals' , xlab = 'Fitted Points')
abline(0,0)
```


```{r finalwtmodel}
fit <- lm(mtcars$mpg ~ mtcars$wt)
res <- residuals(fit)
fits <- fitted(fit)
plot(res ~ fits, main = 'Residuals at Fitted Points For Weight Only', ylab = 'Residuals' , xlab = 'Fitted Points')
abline(0,0)
```



If we look at the mean and standard deviation of the residuals for each of these models.
```{r}
fit <- lm(mtcars$mpg ~ mtcars$transtype + mtcars$wt)
res <- residuals(fit)
mean(res)
sd(res)
```
versus
```{r}
fit <- lm(mtcars$mpg ~ mtcars$wt)
res <- residuals(fit)
mean(res)
sd(res)
```
We see very little difference.

Also if we look at the summary statistics of this model, the t and p values tend to indicate little or no effect is caused on MPG by the type of transmission. However the weight of the vehicle appears to be the driving force.

```{r summarystats}
summary(lm(mtcars$mpg ~ mtcars$transtype + mtcars$wt))
```

## Other Models and Plots

The plots below are each fitted with a linear model. Each shows a relationship to MPG and are also closely related to the weight of the vehicle.

```{r eda4, echo=FALSE}
plot(mtcars$disp, mtcars$mpg, xlab = 'Displacement', ylab = 'MPG' , main = 'Displacement vs MPG')
abline(lm(mtcars$mpg ~ mtcars$disp))
```

```{r eda5, echo=FALSE}
plot(mtcars$hp, mtcars$mpg, xlab = 'Horsepower', ylab = 'MPG' , main = 'Horsepower vs MPG')
abline(lm(mtcars$mpg ~ mtcars$disp))

```{r eda6, echo=FALSE}
plot(mtcars$drat, mtcars$mpg, xlab = 'Rear axle ratio', ylab = 'MPG' , main = 'Rear axle ratio vs MPG')
abline(lm(mtcars$mpg ~ mtcars$drat))
```

Note that the `echo = FALSE` parameter was added to some code chunks to prevent printing of the R code that generated the plot.
